### 2017-05-22 sample-sigconf.tex
Pindah Git

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El
  Paso, Texas USA} 
\acmYear{1997}
\copyrightyear{2016}

\acmPrice{15.00}

\begin{document}

\title{SIG Proceedings Paper in LaTeX Format}
\titlenote{Produces the permission block, and
  copyright information}
\subtitle{Extended Abstract}
\subtitlenote{The full version of the author's guide is available as
  \texttt{acmart.pdf} document}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{B. Trovato et al.}

\begin{abstract}
The tremendous amount of user-generated contents, such as the questions about symptoms and diseases being asked by a user in online forums, or blog posts about dietary and fitness, are potential sources for enriching knowledge in medical document retrieval and health-related text mining applications. Understanding those various contents is simply not a trivial task. Reducing the complex characteristics of contents, including but not limited to linguistic variations, into structured information can help the text interpretation process. A common approach is to extract medical terms from each document using available tools, such as medical entity recognizer. However, we argue that there are many missing pieces of information when we rely only on medical entities or keywords as main information. Hence, we use keyphrases extraction techniques to extract more important information about the concept of the document. Previous works in keyphrases extraction mostly focused on a general domain, long and formal written documents. These methods perform poorly when being applied to user-generated contents and medical documents. In this work, we propose deep bi-directional recurrent network and convolutional neural network as a model for extracting keyphrases. Furthermore, we leveraged word embedding, medical concepts from a knowledge base, and linguistic components as our features. The proposed algorithm achieved F1 score of 82.41\%. This result outperforms state of the art system for extracting keyphrases using a statistical model.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

\keywords{ACM proceedings, \LaTeX, text tagging}

\maketitle

\input{samplebody-conf}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sigproc}

\end{document}