### 2017-05-24 (old)paper-content.tex
Masukin Ablation tp jadi kebanyakan hiks

\section{Introduction}
There are a tremendous amount of user-generated contents about medical information through the internet. As a result, there is a growing need for enriching medical-related knowledge, for which it can be very valuable for several tasks, such as medical document retrieval, medical question answering, and health-related text mining applications \cite{cao2010automatically}. However, understanding those various health-related contents, especially from an online forum, is simply not a trivial task since they are written in free text format (i.e. unstructured format), and often prone to grammatical and typographical glitches. Therefore, reducing the complex characteristics of the textual contents, including but not limited to linguistic variations, into structured information could be very beneficial for the text interpretation process.

One possible solution is to extract medical terms that represent the document they belong to. The common approach is harnessing medical entity recognizer. However, we argue that there are many missing pieces of information when we merely rely on medical entities or terminologies as our main information for representing a particular document. Hence, we leverage keyphrase to cover more valuable information. Keyphrases are usually selected phrases or clauses that can capture the main topic of a given document \cite{turney2000learning}. They can provide readers with highly valuable and representative information, such that looking at keyphrases is sufficient to understand the whole body of a document.

In this paper, we only focus the task of keyphrase extraction on Indonesian medical questions from online forums, such as alodokter\footnote{www.alodokter.com}, health detik \footnote{health.detik.com}, dokter.id \footnote{www.dokter.id}, klikdokter \footnote{www.klikdokter.com}. We present a new keyphrase extraction techniques to extract more important information about the patient intentions and needs, which expressed inside Indonesian medical questions. Furthermore, we also propose a new definition of keyphrases, which is based on our observation of health-related and user-generated questions. Actually, our final goal is to develop medical question answering systems to assist physicians or people obtaining health information. The work presented in this paper is absolutely in line with our goal since important keyphrases of a medical question can be used to formulate a query for the passage or question retrieval subsystem \cite{gong2009improving}. The following example shows the results of keyphrase extraction in a medical document:
	\textit{\\Dok , saya sering sekali merasa \textbf{\emph{kram pada perut kanan bagian atas}} disertai \textbf{detak jantung yang cepat} serta \textbf{keringat dingin} dan \textbf{pusing}. Dan saat kambuh seperti itu, \textbf{nafas sangat sakit}. Seminggu yang lalu sampai terkapar karena tidak kuat menahan sakitnya saat bernafas. apakah ada hubungannya dengan \textbf{diet mayo}? terima kasih sblumnya dok.}
	\texttt{(Doc, i frequently feel a \textbf{cramp on upper right stomach} with \textbf{fasten heart beat} also \textbf{cold sweat} and \textbf{headache}. And when the symptom occur it's \textbf{very hurt when breathing}. Last week i fainted because cannot hold the pain when i was breathing. is there any connection with \textbf{moyo diet}? thanks before)}

It can be seen on the example that keyphrases can be a word, phrase, or even clause. Furthermore, there are many grammatical errors and abbreviation, such as \texttt{"sy"} for the replacement of the word \texttt{"saya (I)"}. It is also worth to note that keyphrase \texttt{"diet mayo (mayo diet)"} is not part of medical entity categories listed by Wahid \cite{skripsiWahid} and Abid \cite{skripsiKakAbid} research. Thus, extracting keyphrase for getting patient needs is not a trivial task.

In our work, we use similar approach as in Cao et al. \cite{cao2010automatically}, which treats keyphrase extraction as a sequence labeling task. The difference is that, in our experiment, we employed and combined various deep learning architectures, such as convolutional neural networks and bi-directional long short-term memory networks, to exploit high level features between neighboring word positions. To improve the quality of our model, we leverage several new hand-crafted features that can handle our keyphrase extraction problems in medical user-generated content, such as word importance and word stickiness features. The word importance feature is enable to rank words in a document by their importance value. The more important a word is to the content of document, the higher its "keywordness" value would seem to be. In addition, we also propose word stickiness feature to make our model constructs keyphrases

\section{Related Works}
In general, keyphrase extraction methods can be divided into two groups: unsupervised ranking (statistical) approach and supervised machine learning approach. For the unsupervised line of research, keyphrase extraction is formulated as a ranking problem, in which each candidate keyphrase is assigned a score that represents its keywordness value. The well-known research in unsupervised approach are RAKE \cite{rake} which use the ratio of word degree and frequency to rank terms. There is also Mihalcea and Tarau \cite{mihalcea2004textrank} who developed a graph-based approach which treats words as vertices and constructs edge between words using co-occurrence.

Furthermore, this kind of approach does not require training data, which is often very difficult to obtain. On the other hand, supervised machine learning approach requires training data that contains a collection of documents with their labeled keywords. Using this approach, keyphrase extraction is usually treated as a classification or sequence labeling task in the level of words or phrases. The first step of this approach generates candidate keyphrases from a particular document. Finally, every candidate keyphrase in the document will be classified as either a keyphrase or non-keyphrase. The well-known method for this approach is KEA \cite{witten1999kea}, they use machine-learning (i.e., Naive Bayes) for classifying candidate keyphrases. The utilization of neural networks in classifying candidate keyphrases also conducted by \cite{ekpNeuralNetworks}. Recently, a well-known supervised approach for keyphrase extraction is based on sequence labeling problem (\cite{zhang2008automatic}, \cite{cao2010automatically}, \cite{zhang2016keyphrase}). The assumption behind this model is that the decision on whether a particular word serves as a keyword is affected by the information from its neighboring word positions.

Unfortunately, there are limited works regarding the task of keyphrase extraction in medical domain, especially in Indonesian language. \cite{ekpMedicalDocumentHybrid} use a hybrid medical knowledge base and statistical approach to extract keyphrases from medical articles. They use stopwords to split candidate keyphrases, then candidates are ranked with two aspects: PF-IDF (Phrase Frequency * Inverse Document Frequency) and domain knowledge which is extracted from medical article. In terms of user-generated medical content, \cite{cao2010automatically} use Conditional Random Fields (CRFs) for extracting keywords from medical questions in online health forum. They use information, such as word location and length as features in their experiments.

\section{METHODOLOGY}
In this work, we view keyphrase extraction problem as a sequence labeling task. That is, given a medical question containing $N$ words $w = (w_1, w_2,...,w_N)$, we want to find the best sequence pf labels $y = (y_1, y_2,..., y_N)$, in which each label is determined using probabilities $P(y_i|w_{i-l},...,w_{i+l},y_{i-l},...,y_{i+l})$ where $l$ is a small number.

Because we have a small dataset, our model would have a small learning material and an end-to-end learning approach will not work well. As a result, we leveraged nine handcrafted features that can help our model to characterize the sequence of keyphrases.

\textbf{Word Embedding.} In the experiments, we use word embedding as input to the neural network. A skip-gram model \cite{mikolov2013wordembed} was used to generate these 128-dimensional vectors for documents in evaluation data. The dimension of the vector was chosen based on \cite{skripsiWahid} experiment. The word embeddings we used in this work were trained using a document that we gathered from Indonesia online forum, medical article, and medical question answering forum.

\textbf{Medical Dictionary.} In our experiment, we leverage a medical knowledge base feature using a medical dictionary. Our medical dictionary contains a list of symptom, disease, treatment, and drug in the Indonesian language. The rationale behind this feature is intentions and needs of patients is more focused on their disease, symptom, treatment, and drug. Based on this feature, we classified words by their appearance in the medical dictionary. The dictionary was built by combining disease, symptom, treatment, and drug dictionaries from \cite{skripsiKakAbid} research. The original dictionaries contained disease, symptom, treatment, and drug separately in each dictionary. We represent this feature as one-hot-vector.

\textbf{Word Length.} This feature represents the length of each word (i.e., the number of characters in every word). This feature is based on Cao et al. \cite{cao2010automatically} research on keyphrase extraction in medical question who added word length as a feature because domain-specific words (e.g. "tuberculosis") tend to be lengthy when compared to common English words, and there is a correlation between the length of a word and its IDF value \cite{cao2010automatically}.

\textbf{Word Position.} We also adapted word position as feature based on Cao et al. \cite{cao2010automatically} research. Based on their observation, an important term sometimes appears toward the end of a clinical question. For example, "medicine for bell's palsy" appear towards the end of a question "what is the best medicine for bell's palsy?".

\textbf{POS-tag.} POS-tag of words was also added as our feature. POS-tag may give model grammatical information and a better understanding of ambiguous words. By our observation, many keyphrases have a common POS pattern. We use Stanford Part-Of-Speech Tagger which pre-trained by Dinakaramani \cite{dinakaramani2014designing}. We represent our POS tag feature as one-hot-vector.

\textbf{Medical Entity.} We use medical entity which annotated by \cite{skripsiWahid} using four categories: drug, treatment, symptom, disease. By our observation, medical entities are sometimes part of keyphrases. Furthermore, medical entities can provide more information about drug or disease which not available in training data or a medical dictionary. We also use one-hot-vector to represent this feature.

\textbf{Abbreviation and Acronym.} We also identified words by their appearance in abbreviation or acronym dictionary gathered by \cite{skripsiKakAbid}. We observed that important word may not be shortened by the users, such as "cancer", "flu", "bell's palsy". One-hot-vector is used to classify whether a word is an abbreviation/acronym or not. For example, the word \texttt{mengapa (why)} sometimes is abbreviated into \texttt{mngp} by some patients.

\textbf{Word Importance.} The purpose of this feature is to rank words in a document by their importance. To extract this feature, we adapted a method from TextRank \cite{mihalcea2004textrank} algorithm for ranking words in a document. They represent words as vertices in the graph and the distance between words as edges. However, in building an undirected graph, we use a word similarity by using our pre-trained word embedding model as a weight for edges. Two words have a connected edge if their similarity is not negative. Moreover, we use a modified PageRank \cite{page1999pagerank} algorithm that consider edges weight in the calculation. Formally, let $G = (V, E)$ be an undirected graph with the set of vertices V and set of edges E, where E is a subset of $VxV$. For a given vertex $V_i$ let $In(V_i)$ be the set of vertices that point to it (predecessors) and let $Out(V_i)$ be the set of vertices that vertex $V_i$ points to (successors). The modified PageRank formula that proposed by \cite{mihalcea2004textrank} can be seen in Formula \ref{eq:pagerank}.
\begin{equation}\label{eq:pagerank}
WS(V_{i})=(1-d) + d * \sum_{V_{i} \in In(V_{i})} \frac{w_{ji}}{\sum_{V_{i} \in Out(V_{j})}}w_{jk}
\end{equation}

\textbf{Word Stickiness.} As a keyphrase, a sequence must be in a valid order. There is some noise such as misuse punctuation by users. For example, "Saya mengalami pusing pada jidat sakit perut dan pandangan kabur (I am having a headache on forehead stomachache and blurred vision)", there is no comma between "jidat (forehead)", "sakit perut (stomachache)" and "dan (and)". Our model may mistake the sequence "(jidat sakit perut)forehead stomachache" as a keyphrase. Based on that problem, we propose a feature that may capture how likely of a given word is occurred together with the word before and after. We use Pointwise Mutual Information (PMI) of a bigram probability to capture the likeness. This feature represented as two-dimensional vector $v = [x1, x2]$, with $x_1$ is the PMI with the word before and $x_2$ is the PMI with the word after. Formula \ref{eq:pmi} is the formula for calculating the PMI value. In that formula, $p(x)$ is the occurrence probability of word $x$, $p(y)$ is the occurrence probability of word $y$, and $p(x, y)$ is the probability of word $x$ and $y$ co-occur together.
\begin{equation}\label{eq:pmi}
PMI(x,y)=log(\frac{P(x, y)}{P(x).P(y)})
\end{equation}
For example, the word \texttt{cancer} in sentence "How to prevent cancer doc ?", the feature value is $f_s(cancer) = [0.56, 0.1]$ where $f_s$ is feature stickiness feature function. It is worth to note that the word \texttt{cancer} is rarely co-occur with the word \texttt{doc}.

We process the sequences word by word. All Features are extracted and concatenated into one vector from every word as the input representation of the model. The output of our model are labels which describe whether a word is either part of keyphrase or not.

We introduce Convolutional Neural Networks (CNNs) to exploit high-level features between neighboring word positions. Based on our observation, a keyphrase is. For example, "Doc, I have a frequent back pain. What happen?". The keyphrase for the previous example is "frequent back pain". In order to know that "back" is part of keyphrase, we need information from the word "frequent" which give an intensity of a symptom and "pain" which refer to "back pain" as the main term. With CNNs, we argue that our model can capture the surrounding information of a word. The output of the CNNs layer will be fed into our next layer which is B-LSTM.

To get a better inference, information from the past and future of sequences can be integrated. This approach is proven effective in sequence labeling task such as semantic role labeling \cite{SMRzhou2015end} and named entity recognition \cite{ma2016end}. Therefore, we utilized bi-directional LSTM (B-LSTM) for extracting structural knowledge by processing sequences both forward and backward. We build up to two layers of B-LSTM.

In order to perform the sequence tagging task, we build a fully connected layer in the top of our model. This layer will classify the output of our model with softmax activation function.

We also experimented on customizing the way we input the features underneath our aforementioned model. Instead of concatenating all features into one vector, which can be seen in Formula \ref{eq:concat}, we tried to give weight for every feature we have. 
\begin{equation}\label{eq:concat}
Z =  concatenate(F_{1}, F_{2}, ..F_{9})
\end{equation}
We argue that each feature has different contribution to the model. In order to do so, we create a new layer underneath our model to do the weighting scenario. The following equation to weight all features can be seen in below:
\begin{equation}
Z =  tanh(W _{1}.F_{1} + W_{2}.F_{2} + .. + W_{n}.F_{9})
\end{equation}

Where $Z$ is the input representation,  $W_{i}$ are the weight for each feature, and $F_{i}$ are the extracted feature in vector form.

\section{Evaluations and Results}
\subsection{Data Collection}
To develop training data, we collected question-answer pairs which previously crawled by Abid \cite{skripsiKakAbid} and Wahid \cite{skripsiWahid} from Indonesian question-answering medical forum. Then, we selected 416 distinct question-answer pairs and removed the answer section for every pair because it's not relevant for obtaining patient intentions and needs. Finally, we manually annotated those questions. The statistical information of our data can be seen in Table \ref{tab:descriptive_stats}.

\begin{table}
	\caption{Statistical information of dataset. W, K, $\bar{N}_{w}$, $\bar{N}_{k}$ are the total of words, number of keyphrases, average number of words, and average number of keyphrases in each question, respectively.}
	\label{tab:descriptive_stats}
	\begin{tabular}{lcccc}
		\toprule
		\#questions&W&K&$\bar{N}_{w}$&$\bar{N}_{k}$\\
		\midrule
		416 & 26747  & 64.76 & 1861 & 4.49 \\

		\bottomrule
	\end{tabular}
\end{table}
\subsection{Evaluation}
For our experiments, we employed 10-cross-validation on our dataset. Evaluation method that we use in our work is using partial evaluation to get precision, recall, F1-measure for evaluation metrics. This evaluation is used because we treat keyphrases as sequences of words. Thus we can use partial evaluation which is used to evaluate Named Entity Recognition system \cite{seki2003probabilistic}.

\subsection{Results}
\begin{table}
	\caption{Model Scenario Result}
	\label{tab:model_scenario}
	\begin{tabular}{lccc}
		\toprule
		Models&Precision&Recall&F-Measure\\
		\midrule
		RAKE & 43.24\% & 68.19\% & 52.90\% \\

		CRF & 63.44\% & 64.94\% & 62.52\% \\

		B-LSTM & 81.88\% & \textbf{83.05\%} & 82.37\% \\

		CNN-B-LSTM & \textbf{82.00\%} & 82.99\% & \textbf{82.41\%} \\

		W-CNN-B-LSTM & 81.08\% & 81.18\% & 81.01\% \\
		\bottomrule
	\end{tabular}
\end{table}

For comparison, we implemented RAKE \cite{rake} and CRF \cite{zhang2008automatic}, which use unsupervised and supervised approach. For CRF, we use the same features proposed above. We also conducted an experiment to proof the effectiveness of our CNN layer by comparing standard B-LSTM and B-LSTM with CNNs layer(CNN-B-LSTM). Furthermore, we also compared our model with or without weighting layer (W-CNN-B-LSTM).

As we can see in Table \ref{tab:model_scenario}, RAKE and CRF underperformed on Indonesian user-generated medical question since RAKE were specially devised for formal text. For CRF, our observation from the result provided by CRF, the method failed to predict long sequences as a keyphrase.

Our CNNs layer also shows improvement in precision and F1 score. From the result above, we suggest that CNNs layer can extract more features implicitly from neighboring words. However, our weighting layer failed to provide some improvement.

\subsection{Other Analysis}
We performed an analysis to the aforementioned model, the result shown there are several errors. Some of the error are uncompleted keyphrases. For example "headache on (pusing pada)", "medicine for (obat untuk)". Our model also failed to capture complex medicine name and disease. This is due to our small data set and not all medicine and disease are captured in our knowledge base. Moreover, there are also special case for keyphrases like "treatment for headache and sore throat (penyembuhan untuk sakit kepala dan tenggorokan)", that keyphrase need to be captured as whole because "sore throat" have a connection with the word "treatment". But our model failed to capture that connection, it extracted "treatment for headache" and "sore throat" separately.

\section{Conclusion}
We have proposed a model to address the task of keyphrase extraction on Indonesian user-generated medical domain. Extracting information patient intentions and needs on user-generated medical contents is not a trivial task due to the fact that the content is usually short, contain many medical terms, and written in an unstructured format, as opposed to formal text. Our model is based on sequence labeling task that employs deep learning approach using convolutional neural networks and bi-directional lstm. Several handcrafted features were proposed, including word importance to detect important word in a document and word stickiness for handling unstructured writing. Although our model outperforms state of the art methods for keyphrase extraction, further improvements are needed especially in evaluation method which cannot cover the whole problem.